---
title: "MC technical replicates"
author: ".."
date: "12 oktober 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(qpcrviia7)
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(Hmisc)
```

## Import and prepare data
```{r}
#Import data
qpcr <- read_qpcr("Data/2017-10-05 FAP-C2C12-Muscle qPCR validation MC.xls")
#Import annotation
ann_samples <- read_annotation("Data/Annotations_samples_124839_technical_replicates.xlsx")
ann_genes <- read_annotation("Data/Annotations_primers_124839_technical_replicates.xlsx")
```

## Technical replicates
### List technical replicates above sd of 0.4
```{r}
list_bad_tech_rep(qpcr)
```

###Save by saving this function into qpcr dataframe
```{r}
qpcr <- qc_tech_rep(qpcr)
```


## Merge and tidy
```{r}
join_and_clean_qpcr(qpcr)
```


## Quality control
### Primers and samples
```{r}
par(mfrow=c(1,1))
hist(qpcr$Avg_CT, main = "CT values")

#Plot CT values for samples
ggplot(qpcr, aes(Sample, Avg_CT)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("CT values per sample")

#Plot CT values for primers
ggplot(qpcr, aes(Gene, Avg_CT)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("CT values per primer")
```


### Remove blanc and 
```{r}
qpcr <- remove_blanc(qpcr)
hist(qpcr$CT, main = "CT after removal blanc")
```

### Set min and max CT
```{r}
set_min_max_CT_endog(10, 40)
hist(qpcr$CT, main = "CT after set baseline")
```

### Split housekeeping genes and endogeneous genes
```{r}
split_genes(qpcr)
```


### Endogenous genes
```{r}
#Double meltcurve peaks
list_double_meltcurves(endog)
remove_bad_meltcurves()
```




### Housekeeping genes
```{r}
#QC on housekeeping genes
qc_hkg(hkg)
```

#Remove Rab35
```{r}
hkg <- remove_primer(hkg, "Rpl13a")
```

```{r}
#Take geometric average all housekeeping genes
mean_hkg <- calculate_mean_hkg(hkg)
qc_mean_hkg(hkg)
mean_hkg$CT_avg_hkg <- remove_outliers(mean_hkg$CT_avg_hkg)
```

## Calculate Delta CT
```{r}
calculate_DCT(endog, mean_hkg, cols = c("Cell_type", "Condition", "Mouse"))
```

## Plot Delta CT
```{r}
theme_set(theme_few())
plot_scatter(endog, x.var = "Cell_type", y.var = "DCt", col.var = "Condition") +
  labs(title = "Scatterplot all genes", y = "Delta CT", x = "Cell type")
plot_scatter_per_gene(endog, "Dcn", x.var = "Cell_type", y.var = "DCt", col.var = "Condition") +
  labs(title = "Scatterplot Decorin", y = "Delta CT", x = "Cell type")
plot_bar(endog, x.var = "Cell_type", y.var = "DCt", fill.var = "Condition") +
  labs(title = "Barplot all genes", y = "Delta CT", x = "Cell type")
plot_bar_per_gene(endog, "Dcn", "Cell_type", y.var = "DCt", fill.var = "Condition") +
  labs(title = "Barplot Decorin", y = "Delta CT", x = "Cell type")
plot_pointrange(endog, x.var = "Cell_type", y.var = "DCt", col.var = "Condition") +
  labs(title = "Pointrange all genes", y = "Delta CT", x = "Cell type")
plot_pointrange_per_gene(endog, "Dcn", x.var = "Cell_type", y.var = "DCt", col.var = "Condition") +
  labs(title = "Pointrange Decorin", y = "Delta CT", x = "Cell type")
```

## Plot relative expression
```{r}
plot_scatter(endog, x.var = "Cell_type", y.var = "rel_expr", col.var = "Condition") +
  labs(title = "Scatterplot all genes", y = "Relative expression", x = "Cell type")
plot_pointrange(endog, x.var = "Cell_type", y.var = "rel_expr", col.var = "Condition") +
  labs(title = "Pointrange all genes", y = "Relative expression", x = "Cell type")
plot_bar_per_gene(endog, "Dcn", x.var = "Cell_type", y.var = "rel_expr", fill.var = "Condition") +
  labs(title = "Decorin", y = "Relative expression", x = "Cell type")
```
